---
title: "Exploring 10 years of daily publications of the Luxembourguish newspaper, *L'Union*"
output: 
  flexdashboard::flex_dashboard:
    theme: yeti
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(tidyverse)
library(textrank)
library(tidytext)
library(udpipe)
library(plotly)
library(ggthemes)
```

Sidebar {.sidebar}
=====================================

```{r}
dateInput('date2',
      label = paste('Select date'),
      value = as.character(as.Date("1860-11-14")),
      min = as.Date("1860-11-12"), max = as.Date("1869-12-31"),
      format = "yyyy/mm/dd",
      startview = 'year', language = 'en-GB', weekstart = 1
    )

selectInput(inputId = "tf_df_words", 
            label = "Select number of unique words for tf-idf", 
            choices = seq(1:10),
            selected = 5)

selectInput(inputId = "textrank_n_sentences", 
            label = "Select the number of sentences for the summary of the article", 
            choices = seq(1:20), 
            selected = 5)
```

*The BnL has digitised over 800.000 pages of Luxembourg newspapers. From those, more than 700.000 
pages have rich metadata using international XML standards such as METS and ALTO. 
Multiple datasets are available for download. Each one is of different size and contains different
newspapers. All the digitised material can also be found on our search platform a-z.lu 
(Make sure to filter by “eluxemburgensia”). All datasets contain XML (METS + ALTO), PDF, original 
TIFF and PNG files for every newspaper issue.* 
Source: https://data.bnl.lu/data/historical-newspapers/

This Shiny app allows you to get summaries of the 10 years of daily issues of the "L'Union" newspaper.
In the first tab, a simple word frequency per article is shown, using the tf-idf method. In the 
second tab, summary sentences have been extracted using the `{textrank}` package.


Word frequency per article
===================================== 
Row
-----------------------------------------------------------------------

### Note: there might be days without any publication. In case of an error, select another date.
    
```{r}
path_tf_idf <- reactive({
    paste0("https://raw.githubusercontent.com/b-rodrigues/newspapers_shinyapp/master/tf_idf_data/newspaper_lunion_", as.character(input$date2), "_tf_idf_data.rds")
})

dfInput <- reactive({
        read_rds(url(path_tf_idf())) %>%
        top_n(as.numeric(input$tf_df_words), tf_idf) %>%
        mutate(word = reorder(lemma, tf_idf)) 
})

renderPlotly({
    df_tf_idf <- dfInput()

    p1 <- ggplot(df_tf_idf,
                 aes(word, tf_idf)) +
                 geom_col(show.legend = FALSE, fill = "#82518c") +
                 labs(x = NULL, y = "tf-doc_idf") +
                 facet_wrap(~doc_id, ncol = 2, scales = "free") +
                 coord_flip() +
                 theme_dark()

    ggplotly(p1)
})
```

Summary of articles {data-orientation=rows}
===================================== 
Row 
-----------------------------------------------------------------------

### The sentence in bold is the title of the article. You can show more sentences in the summary by using the input in the sidebar.
    
```{r}

print_summary_textrank <- function(doc_id, summary, n_sentences){
    htmltools::HTML(paste0("<b>", doc_id, "</b>"), paste("<p>", summary(summary, n_sentences), sep = "", collapse = "<br/>"), "</p>")
}

path_textrank <- reactive({
    paste0("https://raw.githubusercontent.com/b-rodrigues/newspapers_shinyapp/master/textrank_data/newspaper_lunion_", as.character(input$date2), "_textrank_data.rds")
})

dfInput2 <- reactive({
        read_rds(url(path_textrank()))
})

renderUI({
    df_textrank <- dfInput2()

    
df_textrank <- df_textrank %>% 
    mutate(to_print = map2(doc_id, summary, print_summary_textrank, n_sentences = as.numeric(input$textrank_n_sentences)))

df_textrank$to_print
})
```
